# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
setwd("~/Coursera")
data <- read.csv('activity.csv')
data$date <- as.Date(data$date,'%Y-%m-%d')
```

## What is mean total number of steps taken per day?
Below is the histogram of the total number of steps taken each day:
```{r}
new_data <- na.omit(data)
sum_step <- aggregate(new_data$steps, by=list(new_data$date),sum)
colnames(sum_step) <- c('date','sum_steps')
library(ggplot2)
qplot(sum_steps, data=sum_step,geom='histogram',binwidth=1000,fill='red')+theme(legend.position='none')
```
Here is the summarized table of mean and median total number of steps taken per day:
```{r,results="asis"}
library(plyr)
library(xtable)
mm_steps <- ddply(new_data,~date,summarise,mean_steps=mean(steps),median_steps=median(steps))
xt <- xtable(mm_steps)
print(xt,type="html")
```

## What is the average daily activity pattern?
Below is a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days:
```{r}
aver_step <- aggregate(new_data$steps, by=list(new_data$interval),mean)
colnames(aver_step) <- c('interval','average_steps')
qplot(x=interval,y=average_steps,data=aver_step,geom='line',color ='red')+theme(legend.position='none')
```

Interval that contains the maximum number of steps:
```{r}
aver_step[aver_step$average_steps == max(aver_step$average_steps),1]
```

## Imputing missing values
The total number of missing values in the dataset:
```{r}
sum(is.na(data))
```
Replace all of the missing values in the dataset using the mean for that 5-minute interval

```{r}
data2 <- data
rep_func <- function(table1,table2){
  for (i in 1:nrow(table1)){
    if (is.na(table1[i,1])){
      j=i-as.integer((i-1)/288)*288
      table1[i,1] <- table2[j,2]
    }
  }
  return(table1)
}
rep_data <- rep_func(data2,aver_step)
```
Below is the histogram of the total number of steps taken each day based on the new dataset:
```{r}
new_sum_step <- aggregate(rep_data$steps, by=list(rep_data$date),sum)
colnames(new_sum_step) <- c('date','new_sum_steps')
library(ggplot2)
qplot(new_sum_steps, data=new_sum_step,geom='histogram',binwidth=1000,fill='red')+theme(legend.position='none')
```
Here is the summarized table of mean and median total number of steps taken per day based on the new dataset:

```{r,results="asis"}
new_mm_steps <- ddply(rep_data,~date,summarise,mean_steps=mean(steps),median_steps=median(steps))
rep_xt <- xtable(new_mm_steps)
print(rep_xt,type="html")
```

## Are there differences in activity patterns between weekdays and weekends?
